---
title: "Cocktails"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}

library(readr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(janitor)
library(wesanderson)
library(png)
library(ggimage)
library(cowplot)
library(knitr)
library(ggcharts)
library(ggthemes)

```


```{r}
cocktails <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-26/cocktails.csv')
boston_cocktails <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-26/boston_cocktails.csv')

clean_vec <- function (x, refactor=FALSE) {
  
  require(magrittr, quietly=TRUE)
  
  if (!(is.character(x) || is.factor(x))) return(x)
  
  x_is_factor <- is.factor(x)
  
  old_names <- as.character(x)
  
  new_names <- old_names %>%
    tolower(.)
  
  if (x_is_factor && refactor) factor(new_names) else new_names
  
}

#fixing groups
cocktails <- cocktails %>% 
  mutate(
    alcoholic = fct_recode(factor(alcoholic),
                           "Non alcoholic" = "Non Alcoholic"),
    ingredient = clean_vec(ingredient),
    drink = clean_vec(drink),
    glass = clean_vec(glass)
  )

theme_set(theme_solarized())
```

# The distribution of alcoholic vs non-alcoholic drinks

The results are not surprising - most of the cocktails are alcoholic.

```{r}

cocktails %>% 
  ggplot()+
  geom_bar(aes(alcoholic, fill = alcoholic),
           colour = "black")+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 4))+
  xlab("Type of drink")
```

# Top 25 ingredients

I took the top 25 most common ingredients and graphed them in a circular manner - this was inspired by @JacobHakim2 over at Twitter.

```{r}
top <- cocktails %>% 
  group_by(ingredient) %>% 
  summarise(count = sum(n())) %>% 
  arrange(desc(count)) %>% 
  top_n(25)

top_angle <-  90 - 360 * (c(1:nrow(top)) - 0.5)/nrow(top)

top %>% 
  ggplot(aes(reorder(ingredient, -count), count))+
  geom_bar(position="stack", stat="identity", aes(fill = ingredient),
           show.legend = FALSE, colour = "black") +
  scale_x_discrete()+
  ylim(-180, 100) +
  theme_void()+
  coord_polar(direction = 1,
              clip = "off")+
  geom_text(data=top, aes(x=ingredient, y=count, label=ingredient), 
            color="#000000", family="Calibri",
            alpha= 0.8 , size=3, inherit.aes = FALSE,
            angle = top_angle, hjust = -0.1)+
  theme(panel.background = element_rect( color = "white"))+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = nrow(top)))+
    annotate(geom = "text",
           x=0,y=-175,
           hjust=.5, vjust=-1,
           label= paste("top", nrow(top)),
           size=10, lineheight=.8,
           family="serif",
           color="#100404")+
  annotate(geom = "text",
           x=0,y=-175,
           hjust=.5, vjust=.75,
           label="ingredients",
           size=10, lineheight=.8,
           family="serif",
           color="#100404")

```

# Ingredients in alcoholic vs non-alcoholic drinks

I took the most common ingredients in non alcoholic drinks and then compared their occurrence in both categories.

```{r}

top_non_alc <- cocktails %>% 
  filter(alcoholic %in% "Non alcoholic") %>% 
  group_by(ingredient) %>% 
  summarise(count = sum(n())) %>% 
  arrange(desc(count)) %>% 
  top_n(15)

top_non_alc_ingredients <- top_non_alc$ingredient

cocktails %>% 
  filter(ingredient %in% top_non_alc_ingredients,
           alcoholic %in% c("Alcoholic", "Non alcoholic"))%>% 
  group_by(alcoholic, ingredient) %>% 
  summarise(sum = sum(n()))%>% 
  pyramid_chart(x = ingredient, y = sum, group = alcoholic,
                bar_colors = c("#C0CA1B", "4898DC"),
                xlab = "Ingredient occurence")
```

# Types of drink categories:

"Ordinary Drink" appears to be the most common category in this dataset.

```{r}
categories <- cocktails %>% 
  group_by(category) %>% 
  summarise(count = sum(n()))

categories %>% 
  ggplot(aes(reorder(category, count), count))+
  geom_col(aes(fill = category), 
           show.legend = FALSE, colour = "black")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 11))+
  coord_flip()+
  labs(x = "Category", y = "Occurence")
```

# The most common ingredients in ordinary drinks:

Gin is the clear winner here, followed by vodka and lemon juice.

```{r}
cocktails %>%
  filter(category == "Ordinary Drink") %>% 
  group_by(ingredient) %>% 
  summarise(sum = sum(n())) %>% 
  arrange(desc(sum)) %>% 
  top_n(10) %>% 
  lollipop_chart(x = ingredient, y = sum,
                 line_color = "dark green", 
                 point_color = "dark blue")+
  theme_solarized()
```

# Which category uses the most ingredients on average?

## A table:

```{r}
avg_ingr <- cocktails %>%
  group_by(category) %>% 
  summarise(avg_ingredients = mean(ingredient_number)) %>% 
  arrange(desc(avg_ingredients))

kable(avg_ingr,
      col.names = c("Category", "Average ingredient count"))

```

Looks like it is homemade liqueur.

## A graph:

```{r, fig.cap= "Average ingredient count by category"}
cocktails %>%
  group_by(category) %>% 
  summarise(avg_ingredients = mean(ingredient_number)) %>% 
  ggplot()+
  geom_col(aes(reorder(category, -avg_ingredients), avg_ingredients, fill = category),
           show.legend = FALSE, colour = "black")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(x ="Category", y = "Average ingredient count")+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 11))

```

## Ingredients in homemade liqueur:

I wouldn't have expected to see coffee here! 

```{r, fig.width = 4, fig.height= 8}

top_liq <- cocktails %>% 
  filter(category == "Homemade Liqueur") %>% 
  group_by(ingredient) %>% 
  summarise(count = sum(n()))%>% 
  arrange(desc(count))%>% 
  head(10)

top_liq_ingredient <- top_liq$ingredient

cocktails %>% 
  mutate(ingredient = factor(ifelse(ingredient %in% top_liq_ingredient, ingredient, "other"))) %>% 
  filter(category == "Homemade Liqueur") %>% 
  ggplot()+
  geom_bar(aes(category, fill = ingredient),
           color = "black")+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 11))
  

```

# What are the top 5 most commonly recommended glass types?

If you want to have the most luck at having picked the correct glass - go simple and use a cocktail glass.

```{r}
glass_types <- cocktails %>% 
  group_by(glass) %>% 
  summarise(sum = sum(n())) %>% 
  arrange(desc(sum)) %>% 
  top_n(5)

kable(glass_types)

```

# Glass types and drinks:

I:

a) selected the top 5 glass types, 

b) retrieved the top 10 most popular drink types for said glasses,

c) mutated the drink variable to only contain 11 levels: either one of these top drinks, or "other"

d) grouped by glass and drink type and sorted by occurence in a descending order, then took the top 20 results and

e) plotted everything in a bar chart
 

```{r}

top_glass <- glass_types$glass

top_drinks <- cocktails %>% 
  filter(glass %in% top_glass) %>% 
  group_by(drink) %>% 
  summarise(sum = sum(n())) %>% 
  arrange(desc(sum)) %>% 
  top_n(10)

top_drinks <- top_drinks$drink

cocktails %>% 
  mutate(
    drink = factor(ifelse(drink %in% top_drinks, drink, "other"))
  ) %>% 
  filter(glass %in% top_glass) %>% 
  group_by(glass, drink) %>% 
  summarise(count = sum(n())) %>% 
  arrange(desc(count)) %>% 
  top_n(20) %>% 
  ggplot()+
  geom_bar(aes(glass, fill = drink),
           color = "black")+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 13))+
  theme(axis.text.x = element_text(angle = 90))+
  labs(x = "Type of glass",
       y = "Occurence")

```

# Which ingredients are the most commonly used with gin?

The top 3 would be lemon juice, grenadine and dry vermouth.

```{r}
gin <- cocktails %>% filter(ingredient == "gin")
gin_id <- gin$id_drink

gin_combos <- cocktails %>% 
  filter(id_drink %in% gin_id)

gin_combos %>% 
  filter(ingredient != "gin") %>% 
  group_by(ingredient) %>% 
  summarise(sum = sum(n())) %>% 
  arrange(desc(sum)) %>% 
  top_n(10) %>% 
  lollipop_chart(x = ingredient, y = sum,
                 line_color = "orange", 
                 point_color = "red")+
  theme_solarized()
```
  
# Which drinks have the longest names?

There are certainly some intriguing specimens here.
  
```{r}
  long <- cocktails %>% 
  mutate(name_length = nchar(drink)) %>% 
  group_by(drink) %>% 
  summarise(name_length = mean(name_length)) %>% 
  arrange(desc(name_length)) %>% 
  top_n(10) %>% 
  inner_join(cocktails) %>% 
  select(drink, name_length, id_drink) %>% 
  distinct()

kable(long, col.names = c("Drink name", "Length", "Drink Id"))
```

# Let's plot some of the more interesting cases:

Not as complicated as the names make them sound, huh?

```{r, fig.width= 9, fig.height= 5}

cocktails %>% 
  filter(id_drink %in% c(14029, 16447, 13200)) %>% 
  select(drink, ingredient) %>% 
  ggplot()+
  geom_count(aes(drink, ingredient, color = ingredient),
             show.legend= FALSE)+
  facet_wrap(~drink)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_color_manual(values = wes_palette("Zissou1", type = "continuous", n = 10))

```

# A random sample of drinks containing both gin and lemon juice:

```{r, fig.width= 12, fig.height= 6}
set.seed(899)

gin <- cocktails %>% filter(ingredient == "gin")
gin_id <- gin$id_drink

gin_combos <- cocktails %>% 
  filter(id_drink %in% gin_id)

lemon <- cocktails %>% filter(ingredient == "lemon juice")
lemon_id <- lemon$id_drink

lemon_combos <- cocktails %>% 
  filter(id_drink %in% lemon_id)

both_gin_lemon_juice <- inner_join(lemon_combos, gin_combos)

ids <- sample(both_gin_lemon_juice$id_drink, 4, replace = TRUE)

both_gin_lemon_juice %>% 
  filter(id_drink %in% ids) %>% 
  select(drink, ingredient) %>% 
  ggplot()+
  geom_count(aes(drink, ingredient, color = ingredient),
             show.legend= FALSE)+
  facet_wrap(~drink, nrow = 1)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_color_viridis_d()
```

# Boston cocktails

Let's have a short glance at this dataset. What are the top ingredients and how often do they occur in all categories but the Cocktail Classics?

```{r, fig.width = 15, fig.height= 10}
boston_cocktails <- boston_cocktails %>% 
  mutate(
    category = factor(category),
    measure = factor(measure)
  )

top_bost <- boston_cocktails %>% 
  filter(category != "Cocktail Classics") %>% 
  group_by(ingredient) %>% 
  summarise(count = sum(n())) %>% 
  arrange(desc(count)) %>% 
  head(20)

top_bost_ingredients <- top_bost$ingredient

boston_cocktails %>%
  filter(category != "Cocktail Classics") %>% 
  mutate(ingredient = ifelse(ingredient %in% top_bost_ingredients, ingredient, "Other")) %>% 
  ggplot()+
  geom_bar(aes(category, fill = ingredient),
           color = "black")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 21))+
  coord_flip()


```

# An apology to the Cocktail Classics for having excluded them:

Everybody likes classics, right?

```{r, fig.width = 4, fig.height = 8}
boston_cocktails %>%
  filter(category == "Cocktail Classics") %>% 
  mutate(ingredient = ifelse(ingredient %in% top_bost_ingredients, ingredient, "Other")) %>% 
  ggplot()+
  geom_bar(aes(category, fill = ingredient),
           color = "black")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 17))

```

# A random sample of drinks:

The boston cocktail data set provides a measure for each ingredient, which is handy for anyone wanting to test the drink recipes.

```{r, fig.width= 12, fig.height= 6}

set.seed(222)
boston_sample <- sample(boston_cocktails$row_id, 4, replace = FALSE)

boston_cocktails %>% 
  filter(row_id %in% boston_sample) %>% 
  ggplot(aes(name, ingredient, color = ingredient))+
  geom_label(aes(label = measure),
             show.legend = FALSE)+
  facet_wrap(~name, nrow = 1)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_color_manual(values = wes_palette("Zissou1", type = "continuous", n = 16))
  

```